---
title: "ESM 204 Assignment 2: Water Markets"
author: "Evie Vermeer & Lucas Boyd"
date: "4/12/2022"
output: 
  html_document:
    code_folding: hide
    theme: spacelab
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
library(equatiomatic)
library(kableExtra)
```
# Intro {.tabset .tabset-pills}

## 1. Marginal Abatement

***

```{r}
water <- read_csv(here("data", "Water_Districts.csv")) %>% 
  clean_names() %>% 
  mutate(mc_aggregate = mc_kern + mc_antelope + mc_ventura + mc_mojave)

water_longer <- water %>% 
  pivot_longer(cols = c(3:6), names_to = "district", values_to = "mc") %>% 
  mutate(county = case_when(
    district == "mc_kern" ~ "Kern", 
    district == "mc_antelope" ~ "Antelope", 
    district == "mc_ventura" ~ "Ventura", 
    district == "mc_mojave" ~ "Mojave",
    district == "mc_aggregate" ~ "Aggregate"
  ))

ggplot(data = water_longer, 
       aes(x = reduction, y = mc, color = district)) +
  geom_line(size = 0.3) + 
  geom_smooth(se = FALSE, method = "lm") +
  theme_minimal(14) +
  labs (y = "Marginal cost of abatement ($/acre foot)", 
        x = "Level of reduction (acre feet)") +
  scale_color_manual(values = c("cyan4", "goldenrod4", "firebrick", "forestgreen", "black")) +
  theme(legend.position = c(0.2, 0.7)) +
  scale_y_continuous(limits = c(0, 500))
```

**Figure 1** shows marginal cost of abatement of water use for each district, thick, straight lines represent a model fit to the data using simple linear regression.
```{r}
lm_kern <- lm(mc_kern ~ 0 + reduction, data = water)
lm_ventura <- lm(mc_ventura ~ 0 + reduction, data = water)
lm_antelope <- lm(mc_antelope ~ 0 + reduction, data = water)
lm_mojave <- lm(mc_mojave ~ 0 + reduction, data = water)
lm_aggregate <- lm(mc_aggregate ~ 0 + reduction, data = water)
```

### Linear models
Estimating the marginal abatement cost for each district and the aggregate of all districts. 
```{r}
extract_eq(lm_kern, use_coefs = TRUE)
extract_eq(lm_ventura, use_coefs = TRUE)
extract_eq(lm_antelope, use_coefs = TRUE)
extract_eq(lm_mojave, use_coefs = TRUE)
```

```{r}
slope_kern <- lm_kern$coefficients[1]
slope_ventura <- lm_ventura$coefficients[1]
slope_antelope <- lm_antelope$coefficients[1]
slope_mojave <- lm_mojave$coefficients[1]
slope_aggregate <- 1/slope_kern + 1/slope_ventura + 1/slope_antelope + 1/slope_mojave
```

## 2. Demand Curves

***

```{r}
demand_vector <- c(0:250)

use_kern <- 150
use_mojave <- 140
use_antelope <- 220
use_ventura <- 245
use_aggregate <- 755

demand_df <- data.frame(demand_vector) %>% 
  mutate(mb_kern = slope_kern*(use_kern - demand_vector)) %>% 
  mutate(mb_ventura = slope_ventura*(use_ventura - demand_vector)) %>%  
  mutate(mb_mojave = slope_mojave*(use_mojave - demand_vector)) %>% 
  mutate(mb_antelope = slope_antelope*(use_antelope - demand_vector)) %>%
  pivot_longer(cols = c(2:5), names_to = "district", values_to = "mb") %>% 
  mutate(county = case_when(
    district == "mb_kern" ~ "Kern", 
    district == "mb_antelope" ~ "Antelope", 
    district == "mb_ventura" ~ "Ventura", 
    district == "mb_mojave" ~ "Mojave"
  ))

ggplot(data = demand_df, aes(x = demand_vector, y = mb, color = county)) +
  geom_line(size = 1) +
  theme_minimal(14) +
  scale_y_continuous(limits = c(0, 750)) +
  labs (y = "Marginal benefit of water use ($/acre foot)", 
        x = "Water use (consumption) (acre feet)") +
  scale_color_manual(values = c("cyan4", "goldenrod4", "firebrick", "forestgreen")) +
  theme(legend.position = c(0.8, 0.7)) 
```

**Figure 2** shows marginal willingness to pay for water for each district. 

**The Antelope district** is willing to pay the most for the first acre foot of water, at a price of **$628.68**. 

## 3. Policy interventions

#### Water use goal = 500 AF

***

### A. Cap without trade 

**Table 1** displays the results of putting a market cap of 500 acre feet of water without allowing trading. 
```{r}
reduction_target <- 500

stringency_a <- 1-(reduction_target/use_aggregate) # calculating the fractional reduction for all districts

use <- c(use_kern, use_mojave, use_antelope, use_ventura, use_aggregate)
district <- c("Kern", "Mojave", "Antelope", "Ventura", "Aggregate")
mca_slope <- c(slope_kern, slope_mojave, slope_antelope, slope_ventura, slope_aggregate)

template_df <- data.frame(district, use, mca_slope) %>% 
  filter(district != "Aggregate")

cap_no_trade <- template_df %>% 
  filter(district != "Aggregate") %>% 
  mutate(water_use_new = (use*(reduction_target/use_aggregate))) %>% # calculating water use after the mandated reduction
  mutate(abatement = use - water_use_new) %>% # calculating the total abatement required by each district
  mutate(total_cost = (abatement*(abatement*mca_slope))/2) %>% # area under the MCA curve to calculate total cost to each district for that level of abatement
  mutate(stringency = stringency_a) %>% # calculating stringency by dividing abatement by initial use 
  select(district, abatement, total_cost, stringency)

kable(cap_no_trade, digits = 2, col.names = 
        c("District", "Abatement (AF)", "Cost to District (USD)", "Stringency")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left")

total_cost_a <- round(sum(cap_no_trade$total_cost[1:4]), 0)
```

**Total cost of this policy:** $`r paste(total_cost_a)`

There are no taxes collected under this policy intervention so **tax revenue is zero**. 

***

### B. Tax on water use

```{r}

# calculate the necessary tax by solving for 
tax <- (755-reduction_target)/slope_aggregate

tax_df <- template_df %>% 
  mutate(abatement = tax/mca_slope) %>% # calculating abatement by each district by plugging in the tax and solving for quantity abated for each district
  mutate(total_cost = (abatement*abatement*mca_slope)/2) %>% # calculating total cost by finding the area under the curve for each district at that level of abatement
  mutate(stringency = abatement/use) %>% # calculating stringency by dividing abatement by initial use 
  select(district, abatement, total_cost, stringency)

kable(tax_df, digits = 2, col.names = 
        c("District", "Abatement (AF)", "Cost to District (USD)", "Stringency")) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left")
  
total_cost_b <- round(sum(tax_df$total_cost[1:4]), 0) # calculating the total cost to all districts
tax_rev_b <- round(reduction_target*tax, 0) # calculating tax revenue by multiplying the total water use by the tax

```

**Total cost of this policy:** $`r paste(total_cost_b)`

**Tax Revenue:** $`r paste(tax_rev_b)`

*** 

### C. Cap and trade

```{r}
cap_and_trade <- template_df %>% 
  mutate(water_use_no_trade = (use*(reduction_target/use_aggregate))) %>% # pulling code from (A) to calculate the water use before trading
  mutate(abatement_before_trade = use - water_use_no_trade) %>% # calculating the abatement before trading
  mutate(abatement_after_trade = tax/mca_slope) %>% # calculating abatement by each district by plugging in the tax and solving for quantity abated for each district
  mutate(buy_sell = abatement_before_trade - abatement_after_trade) %>% 
  mutate(total_cost_before_trade = (abatement_before_trade*abatement_before_trade*mca_slope)/2) %>% # calculating total cost by finding the area under the curve for each district at that level of abatement
  mutate(total_cost_after_trade = (abatement_after_trade*abatement_after_trade*mca_slope)/2) %>% # calculating total cost of abatement after trading is allowed
  mutate(total_cost = total_cost_after_trade + buy_sell*tax) %>% # addding the revenue generated from buying and selling water rights
  mutate(stringency = abatement_after_trade/use) %>% # calculating stringency by dividing abatement by initial use 
  select(district, abatement_after_trade, total_cost, stringency)

kable(cap_and_trade, digits = 2, col.names = 
        c("District", "Abatement (AF)", "Cost to District (USD)", "Stringency")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left")

total_cost_c <- round(sum(cap_and_trade$total_cost[1:4]), 0) # calculating the total cost to all districts

```

**Total cost of this policy:** $`r paste(total_cost_c)`

There are no taxes collected under this policy intervention so **tax revenue is zero**. 

***

## 4. Drought

#### Water use goal = 300 AF

***

### A. Cap without trade 

**Table 1** displays the results of putting a market cap of 500 acre feet of water without allowing trading. 
```{r}
reduction_target <- 300 # change this number to run for drought

stringency_a <- 1-(reduction_target/use_aggregate) # calculating the fractional reduction for all districts

use <- c(use_kern, use_mojave, use_antelope, use_ventura, use_aggregate)
district <- c("Kern", "Mojave", "Antelope", "Ventura", "Aggregate")
mca_slope <- c(slope_kern, slope_mojave, slope_antelope, slope_ventura, slope_aggregate)

template_df <- data.frame(district, use, mca_slope) %>% 
  filter(district != "Aggregate")

cap_no_trade <- template_df %>% 
  filter(district != "Aggregate") %>% 
  mutate(water_use_new = (use*(reduction_target/use_aggregate))) %>% # calculating water use after the mandated reduction
  mutate(abatement = use - water_use_new) %>% # calculating the total abatement required by each district
  mutate(total_cost = (abatement*(abatement*mca_slope))/2) %>% # area under the MCA curve to calculate total cost to each district for that level of abatement
  mutate(stringency = stringency_a) %>% # calculating stringency by dividing abatement by initial use 
  select(district, abatement, total_cost, stringency)

kable(cap_no_trade, digits = 2, col.names = 
        c("District", "Abatement (AF)", "Cost to District (USD)", "Stringency")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left")

total_cost_a <- round(sum(cap_no_trade$total_cost[1:4]), 0)
```

**Total cost of this policy:** $`r paste(total_cost_a)`

There are no taxes collected under this policy intervention so **tax revenue is zero**. 

***

### B. Tax on water use

```{r}

# calculate the necessary tax by solving for 
tax <- (755-reduction_target)/slope_aggregate

tax_df <- template_df %>% 
  mutate(abatement = tax/mca_slope) %>% # calculating abatement by each district by plugging in the tax and solving for quantity abated for each district
  mutate(total_cost = (abatement*abatement*mca_slope)/2) %>% # calculating total cost by finding the area under the curve for each district at that level of abatement
  mutate(stringency = abatement/use) %>% # calculating stringency by dividing abatement by initial use 
  select(district, abatement, total_cost, stringency)

kable(tax_df, digits = 2, col.names = 
        c("District", "Abatement (AF)", "Cost to District (USD)", "Stringency")) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left")
  
total_cost_b <- round(sum(tax_df$total_cost[1:4]), 0) # calculating the total cost to all districts
tax_rev_b <- round(reduction_target*tax, 0) # calculating tax revenue by multiplying the total water use by the tax

```

**Total cost of this policy:** $`r paste(total_cost_b)`

**Tax Revenue:** $`r paste(tax_rev_b)`

*** 

### C. Cap and trade

```{r}
cap_and_trade <- template_df %>% 
  mutate(water_use_no_trade = (use*(reduction_target/use_aggregate))) %>% # pulling code from (A) to calculate the water use before trading
  mutate(abatement_before_trade = use - water_use_no_trade) %>% # calculating the abatement before trading
  mutate(abatement_after_trade = tax/mca_slope) %>% # calculating abatement by each district by plugging in the tax and solving for quantity abated for each district
  mutate(buy_sell = abatement_before_trade - abatement_after_trade) %>% 
  mutate(total_cost_before_trade = (abatement_before_trade*abatement_before_trade*mca_slope)/2) %>% # calculating total cost by finding the area under the curve for each district at that level of abatement
  mutate(total_cost_after_trade = (abatement_after_trade*abatement_after_trade*mca_slope)/2) %>% # calculating total cost of abatement after trading is allowed
  mutate(total_cost = total_cost_after_trade + buy_sell*tax) %>% # addding the revenue generated from buying and selling water rights
  mutate(stringency = abatement_after_trade/use) %>% # calculating stringency by dividing abatement by initial use 
  select(district, abatement_after_trade, total_cost, stringency)

kable(cap_and_trade, digits = 2, col.names = 
        c("District", "Abatement (AF)", "Cost to District (USD)", "Stringency")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left")

total_cost_c <- round(sum(cap_and_trade$total_cost[1:4]), 0) # calculating the total cost to all districts

```

**Total cost of this policy:** $`r paste(total_cost_c)`

There are no taxes collected under this policy intervention so **tax revenue is zero**. 

***



